---
title: "Reproducible Research week2"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE,message=FALSE}
knitr::opts_chunk$set(echo = TRUE,message=FALSE)
library(dplyr)
library(lattice)
```

Load the dataset in dataframe activity



```{r}
activity<- read.csv("./activity.csv")
```

# 1. What is mean total number of steps taken per day?


```{r}
options(scipen = 1, digits = 2)

# remove NA value
activity1<-subset(activity,!is.na(activity$steps))
# summurie steps number per day
steps<-activity1 %>%
  group_by(date) %>%
  summarize(steps_num=sum(steps))
# plot histogram
hist(steps$steps_num,main="Histogram of daily steps number",xlab="Daily steps number")
# calculate mean value
steps_mean <-mean(steps$steps_num)
steps_median<-median((steps$steps_num))
```

### The mean value of total number steps taken per day is **`r steps_mean`**
### The median value of total number steps taken per day is **`r steps_median`**


# 2. What is the average daily activity pattern?

```{r}
options(scipen = 1, digits = 2)
# remove NA value
activity1<-subset(activity,!is.na(activity$steps))
# summurie steps number per day
intervals<-activity1 %>%
  group_by(interval) %>%
  summarize(steps_avg=mean(steps))
# plot time series
with(intervals,plot(interval,steps_avg,type = "l"),ylab="average number of steps")
# calculate mean value
maks_steps <-intervals[which.max(intervals$steps_avg),1]
```

### The interval with maximum number of average across all the days is **`r maks_steps`**

# 3. Imputing missing values

```{r}
options(scipen = 1, digits = 2)
# calculatenumber of  NA value
na_number<-sum(is.na(activity$steps))
# replece missing values with average number of steps takenin interval
# use ceilling function to get integer numer of steps
temp <-merge(activity,intervals,by="interval")
activity_new <- temp %>%
      mutate(steps = ifelse(is.na(steps),ceiling(steps_avg),steps)) %>% 
      select(steps,date,interval) %>% 
      arrange(date,interval)
# summurie steps number per day for new dataset
steps_new<-activity_new %>%
  group_by(date) %>%
  summarize(steps_num=sum(steps))
# plot histogram
hist(steps_new$steps_num,main="Histogram of daily steps number",xlab="Daily steps number")
# calculate mean value
steps_mean_new <-mean(steps_new$steps_num)
steps_median_new<-median((steps_new$steps_num))
```

### total number of missing values in the dataset is **`r na_number`**
### The mean value of total number steps taken per day is **`r steps_mean_new`**
### The median value of total number steps taken per day is **`r steps_median_new`**
### The mean value differ from the estimates from orginal  **`r steps_mean_new-steps_mean`**
### The median value differ from the estimates from orginal  **`r steps_median_new-steps_median`**
### The total number of steps differ from the estimates from orginal  **`r sum(activity1$steps)-sum(activity_new$steps)`**


# 4. Are there differences in activity patterns between weekdays and weekends?

```{r}
options(scipen = 1, digits = 2)
# add column with name of weekday
activity_new <- activity_new %>%
      mutate(weekday = weekdays(as.Date(date),abbreviate = TRUE))
#change to a factor
activity_new$weekday<-as.factor(activity_new$weekday)
#change levels in factor
levels(activity_new$weekday)= list(weekday= c("Pn","Wt","Œr","Cz","Pt"), weekend = c("So","N"))
intervals_day<-activity_new %>%
  group_by(interval,weekday) %>%
  summarize(steps_avg=mean(steps))
xyplot(steps_avg~interval|weekday,intervals_day,type='l')
```
